---
title: "S-A Correspondence to Age Effect"
author: "Audrey Luo"
date: "2024-02-19"
output: 
  html_document:
    highlight: haddock
    code_folding: hide
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
 

```
 
 
 
 
```{r load files for dti_fa md and SAaxis}

gam.age.dti_fa <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/dti_fa/GAMresults.dti_fa.age_GMfiltered.RData")
gam.age.md <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/md/GAMresults.md.age_GMfiltered.RData")
gam.age.rd <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/rd/GAMresults.rd.age_GMfiltered.RData")
gam.age.ad <- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/ad/GAMresults.ad.age_GMfiltered.RData")
SAaxis <- read.csv("/cbica/projects/luo_wm_dev/SAaxis/SensorimotorAssociation_Axis.fsaverage5.csv", header=T)

write.csv(gam.age.dti_fa, "/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/dti_fa/GAMresults.dti_fa.age_GMfiltered.csv")
write.csv(gam.age.md, "/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/md/GAMresults.md.age_GMfiltered.csv")
write.csv(gam.age.rd, "/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/rd/GAMresults.rd.age_GMfiltered.csv")
write.csv(gam.age.ad, "/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/ad/GAMresults.ad.age_GMfiltered.csv")

 
```

How many vertices have significant age effect after FDR?
```{r check sig age effects}

#length(which(gam.age.dti_fa$s_age.p.value.fdr < 0.05)) # dti_fa: 3278 vertices with significant age effect after FDR correction 
#length(which(gam.age.md$s_age.p.value.fdr < 0.05)) # md: 15028 vertices with significant age effect after FDR correction 

print(paste0("dti_fa: ", length(which(gam.age.dti_fa$s_age.p.value.fdr < 0.05)), " vertices (",  round(length(which(gam.age.dti_fa$s_age.p.value.fdr < 0.05))/20484*100, 2),"%) with significant age effect after FDR correction"))

print(paste0("md: ", length(which(gam.age.md$s_age.p.value.fdr < 0.05)), " vertices (",  round(length(which(gam.age.md$s_age.p.value.fdr < 0.05))/20484*100, 2),"%) with significant age effect after FDR correction"))


print(paste0("ad: ", length(which(gam.age.ad$s_age.p.value.fdr < 0.05)), " vertices (",  round(length(which(gam.age.ad$s_age.p.value.fdr < 0.05))/20484*100, 2),"%) with significant age effect after FDR correction"))

print(paste0("rd: ", length(which(gam.age.rd$s_age.p.value.fdr < 0.05)), " vertices (",  round(length(which(gam.age.rd$s_age.p.value.fdr < 0.05))/20484*100, 2),"%) with significant age effect after FDR correction"))

```


```{r remove vertices that hop gyri in SA axis}

identical(gam.age.dti_fa$element_id[(which(gam.age.dti_fa$s_age.delta.adj.rsq == "NaN"))], gam.age.md$element_id[(which(gam.age.md$s_age.delta.adj.rsq == "NaN"))]) # 370 vertices were fully NaN 
vertices_to_rm <- gam.age.dti_fa$element_id[(which(gam.age.dti_fa$s_age.delta.adj.rsq == "NaN"))] + 1 # get the index of vertices
 
SAaxis_filtered <- SAaxis[-vertices_to_rm,]  
gam.age.dti_fa_filtered <- gam.age.dti_fa[-vertices_to_rm,]
gam.age.md_filtered <- gam.age.md[-vertices_to_rm,]
gam.age.ad_filtered <- gam.age.ad[-vertices_to_rm,]
gam.age.rd_filtered <- gam.age.rd[-vertices_to_rm,]
#nrow(gam.age.md_filtered)
#length(SAaxis_filtered)


gam.age.dti_fa[which(gam.age.dti_fa$s_age.delta.adj.rsq == "NaN"),]
```


```{r remove medial wall}
# remove medial wall
medial_wall <- which(SAaxis_filtered==0) 
SAaxis_filtered <- SAaxis_filtered[-medial_wall]

gam.age.dti_fa_filtered <- gam.age.dti_fa_filtered[-medial_wall,]
gam.age.md_filtered <- gam.age.md_filtered[-medial_wall,]
gam.age.ad_filtered <- gam.age.ad_filtered[-medial_wall,]
gam.age.rd_filtered <- gam.age.rd_filtered[-medial_wall,]

#length(SAaxis_filtered)
#nrow(gam.age.dti_fa_filtered)
```




Correlate with SA axis
```{r correlate w SAaxis}


print("dti_fa")
cor.test(gam.age.dti_fa_filtered$s_age.partial.rsq, SAaxis_filtered, method=c("spearman")) # r = -0.1576228, pspin?
cor.test(gam.age.dti_fa_filtered$s_age.delta.adj.rsq, SAaxis_filtered, method=c("spearman")) # r = -0.1578304 , pspin?

 
 
print("md")
cor.test(gam.age.md_filtered$s_age.partial.rsq, SAaxis_filtered, method=c("spearman")) # r = 0.1699813 , pspin?
cor.test(gam.age.md_filtered$s_age.delta.adj.rsq, SAaxis_filtered, method=c("spearman")) # r = 0.170142  , pspin?



print("ad")
cor.test(gam.age.ad_filtered$s_age.partial.rsq, SAaxis_filtered, method=c("spearman")) # r =   , pspin?
cor.test(gam.age.ad_filtered$s_age.delta.adj.rsq, SAaxis_filtered, method=c("spearman")) # r =    , pspin?


print("rd")
cor.test(gam.age.rd_filtered$s_age.partial.rsq, SAaxis_filtered, method=c("spearman")) # r = 0.1699813 , pspin?
cor.test(gam.age.rd_filtered$s_age.delta.adj.rsq, SAaxis_filtered, method=c("spearman")) # r = 0.170142  , pspin?

```


There are 6 vertices in the fsaverage5 S-A axis that have the same S-A rank assigned. Probably wouldn't have significantly impacted results but more reason to double check wb transform between fsLR and fsaverage5 
```{r troubleshooting SA axis}

## testing

#SAaxis_filtered[which(duplicated(SAaxis_filtered))]
 
#which(SAaxis_filtered=="50729.07")


medial_wall <- which(SAaxis$X0==0) 
SAaxis_noMW <- SAaxis[-medial_wall,]
#SAaxis_noMW[which(duplicated(SAaxis_noMW))]


#which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][1]) # after removing MW, vertices 943 and 10474 have same SA rank
#which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][2]) # after removing MW, vertices 5472 and 11495 have same SA rank
#which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][3]) # after removing MW, vertices 15425 and 15655 have same SA rank


print(paste0("after removing medial wall, vertices ", which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][1])[1], " and ",  which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][1])[2]," have the same SA rank"))


print(paste0("after removing medial wall, vertices ", which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][2])[1], " and ",  which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][2])[2]," have the same SA rank"))


print(paste0("after removing medial wall, vertices ", which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][3])[1], " and ",  which(SAaxis_noMW==SAaxis_noMW[which(duplicated(SAaxis_noMW))][3])[2]," have the same SA rank"))

```


Correlations look very similar without excluding for high GM probability vertices
```{r prelim corr}
```


```{r}

# preliminary S-A axis correlation with age effect

# dti_fa
gam.age.dti_fa_nonfiltered<- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/GAMresults.dti_fa.age.RData")
#write.csv(gam.age.dti_fa_nonfiltered, "/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/GAMresults.dti_fa.age.csv")

SAaxis_nonfiltered <- read.csv("/cbica/projects/luo_wm_dev/SAaxis/SensorimotorAssociation_Axis.fsaverage5.csv", header=T)
medial_wall_nonfiltered <- which(SAaxis_nonfiltered==0)
gam.age.dti_fa_nonfiltered <- gam.age.dti_fa_nonfiltered[-medial_wall_nonfiltered,]
SAaxis_nonfiltered <- SAaxis_nonfiltered$X0[-medial_wall]

# md
gam.age.md_nonfiltered<- readRDS("/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/GAMresults.md.age.RData")
#write.csv(gam.age.md_nonfiltered, "/cbica/projects/luo_wm_dev/output/HCPD/superficialWM/GAM/GAMresults.md.age.csv")

gam.age.md_nonfiltered <- gam.age.md_nonfiltered[-medial_wall_nonfiltered,]
 

print(paste0("dti_fa: ", length(which(gam.age.dti_fa_nonfiltered$s_age.p.value.fdr < 0.05)), " vertices (",  round(length(which(gam.age.dti_fa_nonfiltered$s_age.p.value.fdr < 0.05))/20484*100, 2),"%) with significant age effect after FDR correction"))

print(paste0("md: ", length(which(gam.age.md_nonfiltered$s_age.p.value.fdr < 0.05)), " vertices (",  round(length(which(gam.age.md_nonfiltered$s_age.p.value.fdr < 0.05))/20484*100, 2),"%) with significant age effect after FDR correction"))


print("dti_fa")
cor.test(gam.age.dti_fa_nonfiltered$s_age.partial.rsq, SAaxis_nonfiltered, method=c("spearman")) # r = -0.1576228, pspin?
cor.test(gam.age.dti_fa_nonfiltered$s_age.delta.adj.rsq, SAaxis_nonfiltered, method=c("spearman")) # r = -0.1578304 , pspin?

 

print("md")
cor.test(gam.age.md_nonfiltered$s_age.partial.rsq, SAaxis_nonfiltered, method=c("spearman")) # r = -0.1576228, pspin?
cor.test(gam.age.md_nonfiltered$s_age.delta.adj.rsq, SAaxis_nonfiltered, method=c("spearman")) # r = -0.1578304 , pspin?

```

